Contextuality and the Born rule in fuzzy quantum theories

### 2017-03-28 prop_letter.tex
prop_letter.tex Replace every ket by projectors...

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\cite{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote all eigenstates listed in Table~\ref{table:commonBasis}.
The Kochen-Specker theorem could then be proved by showing there is
no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $ satisfying
equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}\proj{\psi_{i}}}{\begin{gathered}\textrm{mutually orthogonal }\\
\left\{ \ket{\psi_{i}}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}
\end{gathered}
}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}). There must
be a quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}_{24}\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Hence, there exists an interval-valued frame function~$\bar{f}:\mathcal{T}_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}_{24}\left(\proj{\psi}\right)$
satisfying the following conditions because the right-hand side of
each equation forms an orthogonal basis and we can apply Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})\begin{subequations}\label{eq:Kernaghan}
\begin{eqnarray}
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\*
 &  & +\bar{f}(0,0,1,0)+\bar{f}(0,0,0,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\*
 &  & +\bar{f}(0,0,1,1)+\bar{f}(0,0,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,1,0)\nonumber \\*
 &  & +\bar{f}(0,1,0,1)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,0,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(0,1,-1,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,-1,1,1)\nonumber \\*
 &  & +\bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(1,0,1,0)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,-1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(1,0,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\nonumber \\*
 &  & +\bar{f}(0,0,1,1)+\bar{f}(1,-1,0,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,1,-1,0)+\bar{f}(1,0,0,-1)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,-1,-1,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,0,1,-1)+\bar{f}(1,-1,0,0)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,1,-1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,1,0)+\bar{f}(0,1,0,1)\nonumber \\*
 &  & +\bar{f}(1,1,-1,-1)+\bar{f}(1,-1,-1,1)\textrm{ .}
\end{eqnarray}
\end{subequations}If we sum up Eq.~(\ref{eq:Kernaghan}), we will
have
\begin{eqnarray}
\left[11,11\right] & \subseteq & 4\bar{f}(1,0,0,0)+4\bar{f}(1,1,-1,1)\nonumber \\
 &  & +2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{f}\left(\ket{\psi}\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Total}
\end{eqnarray}
where 
\begin{eqnarray}
\mathcal{T}_{18} & = & \mathcal{T}_{24}\backslash\{(1,0,0,0),(1,1,-1,1),(1,-1,1,-1),\nonumber \\
 &  & (1,0,-1,0),(1,0,0,1),(1,1,0,0)\}\textrm{ .}
\end{eqnarray}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in right-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the right-hand
side is a singleton set of an even integer while the left-hand side
is $\left\{ 11\right\} $. Therefore, there is no quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}$ and define $\bar{\mu}_{24}=\left.\bar{\mu}\right|_{\events_{24}}$.

@@

\begin{example}[Four-dimensional quantum three-interval-valued probability
measure]\label{ex:four-dimensional-three-value}Consider the orthonormal
basis~$\left\{ \ket{0},\ket{1},\ket{2},\ket{3}\right\} $ as defined
in Table~\ref{table:commonBasis}. Let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$. Then, we can
define a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{1}$
as follow:\begin{subequations} 
\begin{eqnarray}
\imposs & = & \bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})\nonumber \\*
 & = & \bar{\mu}(\proj{\ps'})\textrm{ ,}\\
\necess & = & \bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})\nonumber \\*
 & = & \bar{\mu}(\mathbb{1}-\proj{\ps'})\textrm{ ,}\\
\unknown & = & \bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}

@@

Second, we will prove that $\bar{\mu}$ cannot correspond to any quantum
probability measure by contradiction. Suppose $\mu(P)\in\bar{\mu}(P)$
for some quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.
We must have 
\begin{equation}
\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0\textrm{ .}\label{eq:probability-zero-on-states}
\end{equation}
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=0}^{N-1}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=0}^{N-1}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=0}^{N-1}q_{j}=1$ and $q_{j}>0$. However, no pure
state~$\ket{\phi}$ satisfies $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$
so that we have $\mu(P)\notin\bar{\mu}(P)$ for all quantum probability
measure~$\mu$.\qed\end{example}

>>>>

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\cite{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote the projectors onto each eigenstate
listed in Table~\ref{table:commonBasis}. The Kochen-Specker theorem
could then be proved by showing there is no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}P_{i}}{\textrm{orthogonal }\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}). There must
be a quantum IVPM~$\bar{\mu}':\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}'\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Because the right-hand side of each equation forms an orthogonal basis,
we can apply Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
and have the following conditions
\begin{eqnarray*}
\necess & \subseteq & \bar{\mu}'\left(P_{00}\right)+\bar{\mu}'\left(P_{01}\right)+\bar{\mu}'\left(P_{10}\right)+\bar{\mu}'\left(P_{11}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{00}\right)+\bar{\mu}'\left(P_{01}\right)+\bar{\mu}'\left(P_{1\ps}\right)+\bar{\mu}'\left(P_{1\ms}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{00}\right)+\bar{\mu}'\left(P_{10}\right)+\bar{\mu}'\left(P_{\ps1}\right)+\bar{\mu}'\left(P_{\ms1}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{00}\right)+\bar{\mu}'\left(P_{11}\right)+\bar{\mu}'\left(P_{\textrm{B}0}\right)+\bar{\mu}'\left(P_{\textrm{B}1}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\textrm{N}3}\right)+\bar{\mu}'\left(P_{\textrm{N}2}\right)+\bar{\mu}'\left(P_{\textrm{N}1}\right)+\bar{\mu}'\left(P_{\textrm{N}0}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\textrm{N}3}\right)+\bar{\mu}'\left(P_{\textrm{N}1}\right)+\bar{\mu}'\left(P_{\ps0}\right)+\bar{\mu}'\left(P_{\ms1}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\textrm{N}2}\right)+\bar{\mu}'\left(P_{\textrm{N}1}\right)+\bar{\mu}'\left(P_{\textrm{B}0}\right)+\bar{\mu}'\left(P_{\textrm{B}3}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\textrm{N}1}\right)+\bar{\mu}'\left(P_{\textrm{N}0}\right)+\bar{\mu}'\left(P_{1\ps}\right)+\bar{\mu}'\left(P_{0\ms}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\textrm{B}1}\right)+\bar{\mu}'\left(P_{\textrm{B}3}\right)+\bar{\mu}'\left(P_{\ps\ps}\right)+\bar{\mu}'\left(P_{\ms\ms}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{1\ms}\right)+\bar{\mu}'\left(P_{0\ms}\right)+\bar{\mu}'\left(P_{\ps\ps}\right)+\bar{\mu}'\left(P_{\ms\ps}\right)\textrm{ ,}\\
\necess & \subseteq & \bar{\mu}'\left(P_{\ps0}\right)+\bar{\mu}'\left(P_{\ps1}\right)+\bar{\mu}'\left(P_{\ms\ps}\right)+\bar{\mu}'\left(P_{\ms\ms}\right)\textrm{ .}
\end{eqnarray*}
If we sum up the preceding equations, we will have
\begin{equation}
\left[11,11\right]\subseteq4\bar{\mu}'\left(P_{00}\right)+4\bar{\mu}'\left(P_{\textrm{N}1}\right)+2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{\mu}'\left(\ket{\psi}\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Total}
\end{equation}
where 
\begin{equation}
\mathcal{T}_{18}=\mathcal{T}_{24}\backslash\{P_{00},P_{\textrm{N}1},P_{\ps\ms},P_{\ms0},P_{\textrm{B}2},P_{0\ps}\}\textrm{ .}
\end{equation}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in right-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the right-hand
side is a singleton set of an even integer while the left-hand side
is $\left\{ 11\right\} $. Therefore, there is no quantum IVPM~$\bar{\mu}':\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}$ and define $\bar{\mu}'=\left.\bar{\mu}\right|_{\events_{24}}$.

@@

\begin{example}[Four-dimensional quantum three-interval-valued probability
measure]\label{ex:four-dimensional-three-value}Consider the orthonormal
basis~$\left\{ \ket{00},\ket{01},\ket{10},\ket{11}\right\} $ as
defined in Table~\ref{table:commonBasis}. Then, we can define a
quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{1}$ as follow:
\begin{eqnarray*}
 &  & \imposs=\bar{\mu}(\mathbb{0})=\bar{\mu}(P_{00})=\bar{\mu}(P_{0\ps})=\bar{\mu}(P_{\ps0})\textrm{ ,}\\
 &  & \necess=\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-P_{00})=\bar{\mu}(\mathbb{1}-P_{0\ps})=\bar{\mu}(\mathbb{1}-P_{\ps0})\textrm{ ,}\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray*}

@@

Second, we will prove that $\bar{\mu}$ cannot correspond to any quantum
probability measure by contradiction. Suppose $\mu(P)\in\bar{\mu}(P)$
for some quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$.
We must have 
\begin{equation}
\mu(P_{00})=\mu(P_{0\ps})=\mu(P_{\ps0})=0\textrm{ .}\label{eq:probability-zero-on-states}
\end{equation}
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=0}^{N-1}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=0}^{N-1}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=0}^{N-1}q_{j}=1$ and $q_{j}>0$. However, no pure
state~$\ket{\phi}$ satisfies $\ip{\phi}{00}=\ip{\phi}{0\ps}=\ip{\phi}{\ps0}=0$
so that we have $\mu(P)\notin\bar{\mu}(P)$ for all quantum probability
measure~$\mu$.\qed\end{example}
